---
title: "Lab 02 - Plastic waste"
author: "Insert your name here"
date: "Insert date here"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

The country of the highest plastic waste per capita is located in North America. Overall, it seems like North America has more countries with comparitively higher plastic waste per capita than other continents.

```{r plastic-waste-continent}
ggplot(
  data = plastic_waste,
  aes(x = plastic_waste_per_cap)
) +
  geom_histogram() +
  facet_wrap(~ continent) +
  labs(x = "Plastic Waste Per Capita", y = "Count of Countries")
```

### Exercise 2

We defined the color and fill of the curves by mapping aesthetics of the plot but defined the alpha level as a characteristic of the plotting geom because we want to have different colors and fills for different continents, while the alpha level to be the same across all continents. Mapping aesthetics allows us to assign unique features to each category, while geom allows us to assign the same feature to all cagtegories.

```{r plastic-waste-density}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.3)
```

### Exercise 3

Remove this text, and add your answer for Exercise 3 here.

```{r plastic-waste-violin}
# insert code here
```

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-mismanaged}
# insert code here
```

```{r plastic-waste-mismanaged-continent}
# insert code here
```

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r recreate-viz}
# insert code here
```


